### 2013-05-21 Problems.tex
Actual questions to answer

\documentclass{article}
\usepackage{amsmath,amscd,amssymb,latexsym, amsfonts}
\usepackage{mathtools}
\usepackage{amsthm}

\theoremstyle{plain}
\newtheorem{iassumption}{Assumption}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{hypothesis}[theorem]{Hypothesis}
\newtheorem{assumption}[theorem]{Assumption}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{question}[theorem]{Question}
\newtheorem{exercise}[theorem]{Exercise}
\newtheorem{statement}[theorem]{Statement}
\newtheorem{example}[theorem]{Example}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}[theorem]{Remark}

\DeclareMathOperator{\Gal}{Gal}
\DeclareMathOperator{\val}{val}
\DeclareMathOperator{\HH}{H}
\DeclareMathOperator{\Ad}{Ad}
\DeclareMathOperator{\Nm}{Nm}
\DeclareMathOperator{\Hom}{Hom}
\DeclareMathOperator{\Spec}{Spec}
\DeclareMathOperator{\Res}{Res}
\DeclareMathOperator{\Fr}{Fr}
\DeclareMathOperator{\Tr}{Tr}
\DeclareMathOperator{\Ind}{Ind}
\DeclareMathOperator{\gen}{gen}

\DeclareMathOperator{\GL}{GL}
\DeclareMathOperator{\PGL}{PGL}
\DeclareMathOperator{\SL}{SL}

\newcommand{\mat}[4]{\left( \begin{array}{cc} {#1} & {#2} \\ {#3} & {#4}
\end{array} \right)}
\newcommand{\TT}{\mathcal{T}}
\newcommand{\C}{\mathcal{C}}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\CC}{\mathbb{C}}
\newcommand{\CCx}{\mathbb{C}^\times}
\newcommand{\OK}{\mathcal{O}_K}
\newcommand{\OKn}{\mathcal{O}_{K_n}}
\newcommand{\pK}{\mathfrak{p}_K}
\newcommand{\pL}{\mathfrak{p}_L}
\newcommand{\OL}{\mathcal{O}_L}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\QQ}{\mathbb{Q}}
\newcommand{\Qp}{\mathbb{Q}_p}
\newcommand{\Gm}{\mathbb{G}_m}
\newcommand{\Lx}{L^\times}

\newcommand{\Weil}{\mathcal{W}}
\newcommand{\WD}{\mathcal{W}'}
\newcommand{\Lpack}{\mathcal{L}}
\newcommand{\Pgen}{P_G^{\gen}}
\newcommand{\bmu}{\boldsymbol\mu}
\newcommand{\mugen}{\bmu^{\gen}}

\newcommand{\la}{\langle}
\newcommand{\ra}{\rangle}

\newcommand{\invlim}[1]{\varprojlim_{#1}}
\newcommand{\Normalizer}[2]{\operatorname{N}_{#2}(#1)}

\begin{document}

\begin{enumerate}

\item For $K$ finite over $\Qp$ and $G$ a unitary group over $K$ splitting over $E$, classify maximal tori in $G$ in terms of \'etale algebras over $E$ with involution.

\begin{conjecture}
They should correspond to \'etale algebras $F / E$ with involution $\sigma$ restricting to the nontrivial involution of $E / K$.  Write $T_F$ for the corresponding torus, which is restriction of scalars of the Norm 1 torus from $F$ to $F^\sigma$.
\end{conjecture}
\item Given $T_F$, which unitary groups does it embed into.

\begin{conjecture}
Given $\kappa \in F^\sigma$, we can construct a hermitian space from the form $\la x, y \ra = \Tr_{F/E}(\kappa \cdot x \sigma(y)$ for $x, y \in F$.  These are precisely the unitary groups into which $T_F$ embeds.
\end{conjecture}

\item What are the invariants of these forms that we've constructed?

\begin{conjecture}
Taking the discriminant of this form should induce a homomorphism from $(F^\sigma)^\times$ to $K^\times / \Nm_{E/K} E^\times$ with kernel $\Nm_{F/F^\sigma} F^\times$.  Have we actually normalized correctly to make this happen.
\end{conjecture}

\item From this description can we read off various properties of the torus: ramified vs unramified, anisotropic, character and cocharacter lattice....

\item Can we give both the cohomological and intrinsic formulation of these questions?

\item What is the global picture?  Are the local-global obstructions?

\item What about $\RR$?

\end{enumerate}

\end{document}