### 2013-05-31 Paper.tex
Moved the 'paper' to a seperate file, should probably be renamed. Otherwise mostly added todo messages about things we should think about.

\documentclass{article}
\usepackage{amsmath,amscd,amssymb,latexsym, amsfonts}
%\usepackage{mathtools}
\usepackage{amsthm}

\theoremstyle{plain}
\newtheorem{iassumption}{Assumption}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{hypothesis}[theorem]{Hypothesis}
\newtheorem{assumption}[theorem]{Assumption}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{question}[theorem]{Question}
\newtheorem{exercise}[theorem]{Exercise}
\newtheorem{statement}[theorem]{Statement}
\newtheorem{example}[theorem]{Example}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{notation}[theorem]{Notation}
\newtheorem{remark}[theorem]{Remark}

\DeclareMathOperator{\Gal}{Gal}
\DeclareMathOperator{\val}{val}
\DeclareMathOperator{\HH}{H}
\DeclareMathOperator{\Ad}{Ad}
\DeclareMathOperator{\Nm}{Nm}
\DeclareMathOperator{\Hom}{Hom}
\DeclareMathOperator{\Spec}{Spec}
\DeclareMathOperator{\Res}{Res}
\DeclareMathOperator{\Fr}{Fr}
\DeclareMathOperator{\Tr}{Tr}
\DeclareMathOperator{\Ind}{Ind}
\DeclareMathOperator{\gen}{gen}
\DeclareMathOperator{\End}{End}

\DeclareMathOperator{\GL}{GL}
\DeclareMathOperator{\PGL}{PGL}
\DeclareMathOperator{\SL}{SL}

\newcommand{\mat}[4]{\left( \begin{array}{cc} {#1} & {#2} \\ {#3} & {#4}
\end{array} \right)}
\newcommand{\TT}{\mathcal{T}}
\newcommand{\C}{\mathcal{C}}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\CC}{\mathbb{C}}
\newcommand{\CCx}{\mathbb{C}^\times}
\newcommand{\OK}{\mathcal{O}_K}
\newcommand{\OKn}{\mathcal{O}_{K_n}}
\newcommand{\pK}{\mathfrak{p}_K}
\newcommand{\pL}{\mathfrak{p}_L}
\newcommand{\OL}{\mathcal{O}_L}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\QQ}{\mathbb{Q}}
\newcommand{\Qp}{\mathbb{Q}_p}
\newcommand{\Gm}{\mathbb{G}_m}
\newcommand{\Lx}{L^\times}

\newcommand{\Weil}{\mathcal{W}}
\newcommand{\WD}{\mathcal{W}'}
\newcommand{\Lpack}{\mathcal{L}}
\newcommand{\Pgen}{P_G^{\gen}}
\newcommand{\bmu}{\boldsymbol\mu}
\newcommand{\mugen}{\bmu^{\gen}}

\newcommand{\la}{\langle}
\newcommand{\ra}{\rangle}

\newcommand{\invlim}[1]{\varprojlim_{#1}}
\newcommand{\Normalizer}[2]{\operatorname{N}_{#2}(#1)}

\newcommand{\TODO}[1]{\textbf{TODO-#1}}

\title{Algebraic Tori in Unitary Groups}
\author{Andrew Fiori \& David Roe}

\begin{document}

\maketitle

\begin{abstract}
We give a characterization of the algebraic tori in unitary groups in terms of the structure of \`etale algebras with involution. We determine relavent properties, such as unramifiedness or anisotropicness, of the tori from the associated algebra. We discuss the implications on the structure of the building for the unitary group. We give a cohomological reinterpretation of the results.
\end{abstract}

\tableofcontents

\section{Introduction}

The problem of giving a classification and/or characterization of algebraic tori in algebraic groups has several sources of motivation coming out of algebraic number theory. In particular it has applications in the the theory of automorphic forms, Shimura varieties and the Langlands conjectures. Moreover, the algebraic tori play a key role in the construction of the building associated to the group, and also in the classification of forms of algebraic groups. Beyond all this the question is natural from a theoretical point of view in the theory of algebraic groups.

\TODO{More Motivational nonsense}

\TODO{References to important literature}

\TODO{Summary of Layout}

\section{Definitions, Notation and Background}

\TODO{} Decide which of the following to define/give notations for, we probably don't need all of it. Some of it can be postponed until the sections where it is used.
\begin{itemize}
\item The setting is a base field $K$ taken to be a $p$-adic field or number field (should we notationally distinguish? I don't currently).
\item We shall denote by $...$ the absolute Galois group of $K$.
\item \'Etale algebras are... typically denoted by $E$.
\item \'Etale algebras with involutions are...
\item Algebraic Tori are... 
\item Hermitian Spaces are ... their invariants over (local fields, global fields) are...
\item Unitary groups are ... their invariants over (local fields, global fields) are...
\item Quadratic Forms are ...their invariants over (local fields, global fields) are...
\item Orthogonal Groups are ...their invariants over (local fields, global fields) are...
\item Connections between orthogonal and unitary groups.
\end{itemize}

\section{Tori and \'Etale Algebras}

Let $T$ be a torus defined over $K$ and $\rho:T_K \rightarrow \GL(V_K)$ a $K$-rational algebraic representation of $T$ (one can and might as well assume $\rho$ is injective).

Associated to this data there exists an \'etale algebra $E$ over $K$, inclusions $\rho'' : E \hookrightarrow \End(V_K)$ and $\rho' : T \rightarrow T_E$ for which $\rho = \rho''\circ\rho'$.

\begin{proposition}
We have the following constructions of $E$:
\begin{enumerate}
\item Let $X$ denote the set of characters of $T$ appearing via the representation of $\rho$ over $\overline{K}$. This is a Galois set for the action of $\Gal(\overline{K}/K)$ hence defines a unique \'etale algebra $E$.
\item Let $C = C_\rho(T)$ be the centralizer of the image of $\rho(T)$ in $\End(V_K)$ (as a scheme over $K$).
           Let $Z = Z(C_\rho(T))$ be the centre of $C$ (as a scheme over $K$).
           Let $E = Z(K)$.
\item If the characters $\chi$ appearing in $X$ are distinguished by their values on $K$ points then $E$ is the $K$-span in $\End(V_K)$ of $\rho(T(K))$.
\end{enumerate}
\end{proposition}
\begin{proof}
That (2) and (3) give \'etale algebras follows from base change to the algebraic closure.
That the algebra constructed by (3) is contained in that constructed by (2) is clear.
The converse can be checked over the algebraic closure, using flatness and dimension considerations.

The equivalence of (1) and (3) follows from the characterization of \'etale algebras by Galois descent.
\end{proof}

\begin{remark}
If the size of $X$ equivalently the $K$-dimension of $E$ equals the $K$ dimension of $V_K$ then $V_K$ is a rank $1$ module over $E$ hence may be (non-canonically) identified with $E$.
\end{remark}

\subsection{Structures on $E$}

The choice of representation $\rho$ determines structures on $E$.

\begin{proposition}
Let $B : V_K \times V_K \rightarrow K$ be a non-degenerate bilinear form preserved by the image of $\rho$ then the adjoint involution on $\End(V_K)$ (relative to $B$) defines an involution $\sigma$ on $E$.
Moreover, $T \rightarrow T_{E\sigma} \rightarrow T_E$ is a factorization of the map $\rho'$.
\end{proposition}
\begin{proof}
Denote by $\sigma$ the adjoint involution on $\End(V_K)$.
We are given that $B(x,y) = B(\rho(t)x,\rho(t)y) = B(x,\sigma(\rho(t))\rho(t)y)$ for all $x$ and $y$. By the non-degeneracy of $B$ this implies $\sigma(\rho(t))\rho(t)$ is the identity and thus $\sigma(\rho(t)) = \rho(t)^{-1}$. Thus, $\sigma$ takes $T$ to $T$. by the naturality of the construction of $E$ above $\sigma$ takes $E$ to $E$.

It is now clear that $\rho(T) \subset T_{E,\sigma}$.
\end{proof}

\begin{proposition}
Let $A$ be an algebra over $K$ and suppose $V_K$ has the structure of an $A$-module.
Suppose $\rho(T)$ embeds into $\End_A(V_K)$ then $E\hookrightarrow \End_A(V_K)$.

Suppose $\rho(T)$ commutes with the action of $A$, then the image of $E$ commutes with the action of $A$.
If moreover the $K$-dimension of $E$ equals that of $V$ this also gives $E$ the structure of an $A$-algebra.
\end{proposition}
\begin{proof}
By naturality of construction if $\rho(T)$ acts by $A$ endomorphisms then so too does $E$.
Likewise if $\rho(T)$ commutes with the action of $A$, then so too does $E$.

Finally, if $K$-dimensions are equal, then by identifying $V_K$ with $E$ we have given $E$ the structure of an $A$-module. Since the $A$ and $E$ multiplication commute, we are giving $E$ the structure of an $A$-algebra.
\end{proof}
\begin{remark}
The case of non-commutative $E$ is mostly relevant to inner forms of symplectic groups.
In what follows we will assume $A$ is a commutative algebra over $K$.
\end{remark}

\begin{proposition}
Let $A$ be a quadratic algebra over $K$. Suppose $V_K$ has the structure of an $A$-module, and $B : V_K \times V_K \rightarrow A$ is a non-degenerate Hermitian form.
Suppose  $\rho(T)$ preserves $B$ then $E\hookrightarrow \End_A(V_K)$, the image of $E$ commutes with the action of $A$,  $E$ has the structure of an $A$-algebra, and the involution on $E$ restricts to that on $A$.
\end{proposition}
\begin{proof}
The only new statement here is that the involution on $E$ restricts to the involution on $A$.
The involution on $E$ was induced by the adjoint from $B$. But the involution on $A$ is by construction the adjoint.
\end{proof}
\begin{remark}
The $K$-algebra $E= A\times A$ has two meaningfully different $A$-algebra structures arising from the inclusions:
\[ a \mapsto (a,a) \qquad \text{and}\qquad a\mapsto (a,\sigma(a)) \]
once we impose an involution on $E$. The two cases are meaningfully distinguished by how the involution restricts to $A$.
\end{remark}

\subsection{Tori in Algebraic Groups}

Let $G$ be an algebraic group over $K$ and suppose $T \hookrightarrow G$.
Then any representation $\rho$ of $G$ restricts to a representation $\rho$ of $T$.

The structures imposed on $E_T$ by $\rho$ are determined more by the representation $\rho$ of $G$ and not directly by $T$. If $\rho$ is injective the consequence is that all tori (at least conjugate to $T$ over $\overline{K}$) in $G$ will have the same structures.

\begin{example}
Let $\rho:G \hookrightarrow \GL(V_K,B)$ be a representation preserving the bilinear form $B$ then for all tori $T$ in $G$, there exists an \'etale algebra with involution $(E,\sigma)$ and and inclusion $T \hookrightarrow T_{E,\sigma}$.

This holds even if $T$ is a split torus in the split unitary group $G=GL_n$ and $\rho$ is the representation corresponding to the standard representation of the unitary group. Ie, if the representation corresponds to the same vector in the $A_n$ root system.
\end{example}

\begin{example}
If the form $B$ above is Hermitian for an action of $A$ on $V_K$ then $E$ is an $A$-algebra with the involution on $E$ restricting to that on $A$.

In the split case the algebra $A$ is $K\oplus K$.
\end{example}

\section{Forms preserved by $T_{E,\sigma}$}

\subsection{Shape of the Forms}

\begin{theorem}
Suppose $B_1,B_2$ be bilinear (or Hermitian) forms on $E$ for which the adjoint maps $Ad_{B_1}$ and $Ad_{B_2}$ induce the same map $\sigma:E\rightarrow E$.
That is $B_i(ex,y) = B_i(x,\sigma(e)y)$, ie, we are in the case where $T_{E,\sigma}$ preserves $B_1$ and $B_2$.
Then, $B_1(x,y) = B_2(\lambda x,y)$.
\end{theorem}
\begin{proof}
Both $B_1$ and $B_2$ induce isomorphism between $E$ and its K (resp. A)-linear dual.
Hence, there exists a unique $\lambda\in E$ such that $B_1(1,y) = B_2(\lambda,y)$ for all $y\in E$.
We have:
\[ B_1(e,x) = B_1(1,\sigma(e)x) = B_2(\lambda,\sigma(e)x) = B_2(\lambda e,x). \]
\end{proof}
\begin{theorem}
If in the above if $B_1$ and $B_2$ are symmetric then $\sigma(\lambda)=\lambda$.

If in the above $B_1$ and $B_2$ are Hermitian then $\sigma(\lambda)=\lambda$.
\end{theorem}
\begin{proof}
We have the following calculation:
\[ B_2(\lambda y,x) = B_1(y,x) = B_1(x,y) = B_2(\lambda x,y) = B_2(x,\sigma(\lambda) y) = B_2(\sigma(\lambda)y,x) \]
Alternatively:
\[ B_2(\lambda y,x) = B_1(y,x) = \overline{B_1(x,y)} = B_2(\lambda x,y) = \overline{B_2(x,\sigma(\lambda) y)} = B_2(\sigma(\lambda)y,x) \]
And this holds for all $x,y$.
Hence, by the uniqueness of $\lambda$ from above, $\lambda=\sigma(\lambda)$.
\end{proof}
\begin{corollary}
All the non-degenerate symmetric bilinear forms on $E$ are of the form:
\[ Tr_{E/K}(\lambda x\sigma(y) \]
All the non-degenerate Hermitan forms on $E$ are of the form:
\[ Tr_{E/A}(\lambda x\sigma(y) \]
where $\lambda \in E^\sigma$ is a unit.
In both cases, the isomorphism class of the form depends only on $\lambda \in (E^\sigma)^\times/N_{E/E^\sigma}(E^\times)$.
\end{corollary}

\subsection{Invariants of Hermitian Forms}

\TODO - in this section we might benefit from renormalizing the form, in my experiance letting $\lambda' = \lambda/f'_z(z)$ where $z\in E^\sigma$ is such that $\sqrt{z}$ primitively generates $E$ over $K$ is a good choice.
(Such $z$ exist outside characteristic $2$ when $E$ is a field, or $E$ is an \'etale algebra and $K$ infinite).

\TODO - probably want a lemma about how, or perhaps explaining invariants of Hermitian forms.
$(d,\delta_{A/K}) $ for all intents and purposes is $d \in K^\times/N_{A/K}(A^\times)$.

\begin{notation}
Given an \'etale algebra with involution $(E,\sigma)$ and $\lambda \in (E^\sigma)^\times$
denote the Hermitian form:
\[ H_{E,\sigma,\lambda} := Tr_{E/A}(\lambda x\sigma(y)) \]
and the symmetric bilinear form:
\[ B_{E,\sigma,\lambda} := Tr_{E/A}(\lambda x\sigma(y)). \]
\end{notation}

\TODO - in all the following the Witt invariant formula might be nicer

\begin{proposition}\label{prop:Hasse}
The Hasse invariant of $B_{E,\sigma,\lambda}$ is:
\[ Cor_{E^\sigma/K}((-\lambda f_z'(z),z)(-1,-1)^{n(n-1)/2}...). \]
\end{proposition}
This is \TODO - cite

\begin{proposition}\label{prop:InvariantRelations}
Let $H$ be any Hermitian form and denote:
\[ B(x,y) = Tr_{A/K}(H(x,y)). \]
Then:
\begin{enumerate}
\item The discriminant of $H$ (and the discriminant of $A$) determines the Hasse invariant of $B$. (The converse holds meaningfully whenever $A$ is not split, and is trivial otherwise.)
\item The discriminant of $A$ determines the discriminant of $B$.
\item The signatures of $H$ (and the infinite ramification of $A$) determines the signatures of $B$. (The converse holds meaningfully if $A$ is ramified, and is trivial otherwise.)
\end{enumerate}
In particular the invariants of $B$ are equivalent to those of $H$.
\end{proposition}
\begin{proof}
\TODO - complete sketch.

By using Proposition \ref{prop:Hasse} one can compute that the Hasse invariant of the form:
\[ \tfrac{1}{2}\Tr_{A/K}(dx_0\sigma(y_0) + \sum x_i\sigma(y_i)) \]
is \TODO - work out exact 
\[ (d,\delta_{A/K})(-2\delta_{A/K},\delta_{A/K})^n(-\delta_{A/K},-\delta_{A/K})^{n(n-1)/2}... \]
Conversely, it follows that:
\[ (d,\delta_{A/K}) = Hasse (...). \]
At all places where $A/K$ is not split, this determines $d$ in $K^\times/N_{A/K}(A^\times)$.
At all places where $A/K$ is split, the norm map is surjective, and the discriminant is meaningless.
\TODO - phrasing that tells us we just determined $d$ globally.

By \TODO - cite the discriminant of the form $B$ is precisely
\[ (-\delta_{A/K})^n. \]

For a real ramified place the signature of $B$ is precisely:
\[ (2s,2r) \]
where $(s,r)$ is the signature of $H$.
For a real split place the signature of $B$ is precisely:
\[ (n,n) \]
whereas, $H$ has no meaningful signature (the algebra $A$ is $\RR\times \RR$) as $-1$ is a norm.

For complex places there are no signatures in either case.
\end{proof}

\begin{proposition}
The discriminant of invariant of $H_{E,\sigma,\lambda}$ is:
\[ TODO \]
Moreover, the map:
\[ (E^\sigma)^\times/N_{E/E^\sigma}(E^\times) \rightarrow K^\times/N_{A/K}(A^\times) \]
is a homomorphism.
\end{proposition}
\begin{proof}
\TODO - complete sketch

Combining the results of Propositions \ref{prop:Hasse} and \ref{prop:InvariantRelations} we find that:
\TODO - details
\[ (d,\delta_{A/K}) =  Cor_{E^\sigma/K}((\lambda,z)) Cor_{E^\sigma/K}((-f'(z),z) (-1,-1)^{n(n-1)/2}...) ... \]
Now, if $\delta_{A/K}$ is a square then so is $z$

Provided $E/E^\sigma$ (equiv $A/K$) not split the formula should tell us non-trivial $d$ come from non-trivial $\lambda$.

Note that  $(E^\sigma)^\times/N_{E/E^\sigma}(E^\times)$ is $( K^\times/N_{A/K}(A^\times))^\ell$ where $\ell$ is the number of field factors of $E$.
\end{proof}

\begin{proposition}
The signature of $H_{E,\sigma,\lambda}$ is:
\[ TODO \]
\end{proposition}

\subsection{Existence of Forms}

\begin{theorem}
Given $H$ a Hermitian form and $(E,\sigma)$ an \'etale algebra with involution (of the second kind) over $A$.Then for all $p$ primes of $K$ there exists $\lambda_p \in E^\sigma$ with:
\[ H_p \simeq H_{E_p,\sigma,\lambda_p} \]
\TODO - I am pretty sure there are no conditions, I haven't fully checked.
\end{theorem}
\begin{proof}
\TODO - complete sketch

non-split Infinite places (use signature proposition)

non-split finite places (use discriminant map proposition)

Split places (trivial)
\end{proof}

\begin{theorem}
\TODO - phrasing

There exists $\lambda_p$ locally everywhere if and only if there exists $\lambda_p$ globally.
\end{theorem}
\begin{proof}
\TODO - complete sketch (also check that this is true).

Argument 1:

The local global obstruction allowing $\lambda$ such that $B_{E,\sigma,\lambda} \simeq B$ does not exist because for any place $p$ where $A/K$ is innert, then for every field factor of $E$ we have $E/E^\sigma$ is innert, thus there is no local global obstruction.

The equivalence of $B$'s and $H$'s should conclude the result.

\bigskip
Argument 2:

Optimistically global class field theory has a name for what we are doing.
We need a $\lambda$ with controlled number of positive/negative embeddings, asside from this we are trying to say that the map:
\[ (E^\sigma)^\times/N_{E/E^\sigma}(E^\times) \rightarrow K^\times/N_{A/K}(A^\times) \]
is surjective.
So in total what we want is surjective when we appropriately restrict to `totally postive' elements. This sounds like a very standard CFT type problem, and since by Argument 1 this result should hold... this result should have been proven by someone else before now.
\end{proof}

\section{Properties of $T_{E,\sigma}$}

\subsection{Anisotropic}

\begin{proposition}
The torus $T_{E,\sigma}$ is anisotropic if all the field factors of $E$ are $\sigma$-stable.
\end{proposition}
\begin{proof}
\TODO-
That the condition is necissary is reasonably clear, sufficiency is less obvious.
\end{proof}

\begin{example}
Consider $E = A \times A$ where $\sigma$ acts on each factor.
This has the structure of an $A$ algebra, where $\sigma$ restricts to $\sigma$, by the inclusion $A\hookrightarrow A\times A$:
\[ a \mapsto (a,a) \]
The torus $T_{E,\sigma}$ is anisotropic.

\bigskip
Consider $E = A \times A$ where $\sigma$ acts to interchange factors.
Give this the structure of an $A$ algebra, where $\sigma$ restricts to $\sigma$, by the inclusion $A\hookrightarrow A\times A$:
\[ a \mapsto (a,\sigma(a)) \]
The torus $T_{E,\sigma}$ is not anisotropic.
\end{example}

\subsection{Unramified}

\begin{proposition}
The torus $T_{E,\sigma}$ is unramified if all the field factors of $E$ are unramified
\end{proposition}
\begin{proof}
\TODO-
\end{proof}

\subsection{``Relatively'' Unramified}

Your special unramifiedness when $A/K$ was ramified.

\section{The Building of the Unitary Group}

\subsection{The reduction of $T_{E,\sigma}$}

\section{Cohomological Reinterpretation}

This is stuff I don't recall super well, I think the following is roughly true, you might recall better than I.

\begin{question}
The set of pairs $(T,U)$ consisting of $T$ a maximal torus (up to rational conjugacy) of $U$ and $U$ a pure inner form of a unitary group are in bijection with:
\[ H^1( \Gal(\overline{K}/K), N_{U_0}(T_0) ) \]
\end{question}
If $T_0$ is split this becomes:
\[ H^1( \Gal(\overline{K}/K), W(U,T_0) ) \]

\begin{question}
The set of pure inner forms of $U$ containing a torus isomorphic to $T_0$ is in bijection with the kernel of the natural map:
\[ H^1( \Gal(\overline{K}/K), N_{U_0}(T_0) ) \rightarrow H^1( \Gal(\overline{K}/K),N_{GL_n}(T_E )) \]
\end{question}

\begin{question}
the set of maximal tori $T$ in $U_0$ (up to rational conjugacy) is in bijection with the kernel of the natural map:
\[  H^1( \Gal(\overline{K}/K), N_{U_0}(T_0) ) \rightarrow H^1( \Gal(\overline{K}/K), U_0) \]
\end{question}

The above should somehow give us that:
\begin{question}
The forms of $T$ contained in $U$ are in bijection with:
\[ H^1(\Gal(\overline{K}/K), W(U_0,T_0)^{\Gal(\overline{K}/K} ) \]
\end{question}
However this could plausibly require that $T_0$ be split or that $U_0$ is quasi-split.

\end{document}