### 2014-03-22 references.tex
Added references file, made some changes

\documentclass{amsart}
\begin{document}

\begin{thebibliography}{9}

\bibitem Lawrence Morris, Some tamely ramified supercuspidal representations of symplectic groups, Proc. London Math. Soc. 63 (1991), no. 3, 519-551.

\bibitem Kazutoshi Kariyama, On Conjugacy Classes of Maximal Tori in Classical Groups.  Journal of Algebra, 125 (1989), 133-149

\bibitem Tignol et. al.  The Book of Involutions.

\bibitem Mark Reeder.  Elliptic Centralizers in Weyl Groups and their Coinvariant Representations.  Representation Theory, 15, (2011), pages 63-111.

\end{document}